---
title: "DAISIE Refactor"
author: "Pedro Santos Neves"
date: "10-02-2023"
output: 
    ioslides_presentation:
        widescreen: true
        logo: logo.png
        css: styles.css
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library("DAISIE")
library(CodeDepends)
```

## Index
* Overview / status
  + Inference, simulation, plotting
  + Compiled/foreign code (C++, FORTRAN)
  + Dependencies
* Improvements
  + Interface
  + Maintainability
  + Style
* Feedback

# Overview / status

## DAISIE at a glance

```{r}
knitr::kable(cloc::cloc_pkg(source = "C:/code/DAISIE"))[1:7]
```

## Inference

* Accessed primarily (but not exclusively) via the `DAISIE_ML_*()` functions
* 5 (6 if redundant `DAISIE_ML()` is counted) functions, common elements in the function tree
* General use of many arguments, defaults prevalent along function trees

## Simulation

* Accessed via the `DAISIE_sim_*()` functions
* 7 (8, if redundant `DAISIE_sim()` is counted) functions, common elements in the function tree

## Plotting

* Accessed via the `DAISIE_plot_*()` functions
* 4 functions, one part of function tree of simulation code

## Acessory functions

* `DAISIE_dataprep()`, `DAISIE_probdist()` etc

## Dependencies

```{r package dependencies2, warning=FALSE, echo=FALSE}
pak::pkg_deps_tree(pkg = "rsetienne/DAISIE", dependencies = TRUE)
```

## Dependencies

```{r package dependencies, echo=FALSE, message=FALSE, warning=FALSE}
depgraph::plot_dependency_graph(
  pkg = "C:/code/DAISIE",
  suggests = FALSE,
  option = "cividis"
)
```

## Dependencies

```{r package dependencies3, warning=FALSE, echo=FALSE}
cran_db <- cranly::clean_CRAN_db()
package_network <- cranly::build_network(cran_db)
dep_table <- cranly::compute_dependence_tree(package_network, "DAISIE")

knitr::kable(dep_table |> dplyr::group_by(generation) |> dplyr::count())
```

## Dependencies

* High number of dependencies
* Some package dependencies themselves have a high number of upstream dependencies
* To maintain the package on CRAN, all dependencies on Imports, Depends and Suggests must be equally available
* Critical dependencies include Rcpp, deSolve, DDD (for ML optimization, long term free-standing package?)
* The lower the number of dependencies, higher reliability

## Compiled/foreign code (C++, FORTRAN)

* Numerical integration of ODE system implemented via R, C++ (using BOOST libraries) and FORTRAN code
* Redundant methods are helpful to compare with each other and as fallbacks when one fails
* Increase maintenance cost. Arguably breaks DRY rule.

# Improvements

## Interface
```
DAISIE_sim_cr(
  time,
  M,
  pars,
  replicates,
  divdepmodel = "CS",
  nonoceanic_pars = c(0, 0),
  num_guilds = NULL,
  prop_type2_pool = NA,
  replicates_apply_type2 = TRUE,
  sample_freq = 25,
  plot_sims = TRUE,
  hyper_pars = create_hyper_pars(d = 0, x = 0),
  area_pars = create_area_pars(max_area = 1, current_area = 1, proportional_peak_t = 0,
    total_island_age = 0, sea_level_amplitude = 0, sea_level_frequency = 0,
    island_gradient_angle = 0),
  cond = 0,
  verbose = TRUE,
  ...
)
```

## Interface

* Standardize argument names: choose and rigidly enforce style guide
* Complex arguments should always be named (lists) at the interface level, and as much as possible in inner functions
* Helper functions for complex arguments are an option (e.g., `create_hyper_pars()`)
  + More space for long form documentation, defaults and ease of testing
* Don't export many currently exported functions

## Maintainability

* Defaults *only* present in user interface
* Deprecate R and FORTAN implementations to a separate package
  + Allows code be kept for tests
  + Same maintenance cost
* Remove dependencies. Consider removing most if not all plotting functions
* Rename ambiguously named functions (i.e., `DAISIE_ML1()`, `DAISIE_ML2()`, etc.)
* Trim down repeated (nearly) identical portions of code (e.g. `DAISIE_ML_*()` functions until `DAISIE_loglik()`)

## Maintainabilty

```{r fig.height=20, fig.width=20, message=FALSE, warning=FALSE}
gg <- makeCallGraph("package:DAISIE")
if(require(Rgraphviz)) {
    gg = layoutGraph(gg, layoutType = "circo")
    graph.par(list(nodes = list(fontsize = 55)))
} 
renderGraph(gg)
```

## Feedback

* Thanks for listening! Any comments?
